---
title: "Assignment 2"
author: "Alex Brown"
date: "5/20/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, messages=FALSE}
#install packages
library(tidyverse)
library(RSQLite)
library(DBI)
```

1. Read the parcels data as-is (all columns; no automatic type conversion) into a parcels_raw tibble.
```{r}

UseCode <- read_delim("UseCodes.csv", delim = "|", , col_types = cols(.default = col_character()))

UseCode

parcels_raw <- read_csv("Santa_Barbara_County_parcels_2011.csv", col_types = cols(.default = col_character()))

parcels_raw

# read in UseCodes and Santa Barbara County parcels (2011) csv's. Removed NA values with col_types command.
```

## Import and Tidy
```{r}

parcels <- transmute(parcels_raw,
    APN,
    Situs1,
    Situs2,
    Acreage = parse_double(parcels_raw$Acreage),
    UseCode,
    NonTaxCode,
    AgPres,
    LandValue = parse_double(parcels_raw$LandValue),
    Net_Impr = parse_double(parcels_raw$Net_Impr),
    Net_AV = parse_double(parcels_raw$Net_AV),
    M_Address1,
    M_Address2)
  
wrote_parcels <- write_delim(parcels,
  "parcels.csv",
  delim = "|",
  na = ""
)

wrote_parcels

# Combining UseCode and Parcels
UseCodeParcels <- parcels %>%
  left_join(UseCode, by = "UseCode")
```

```{r}

con <- dbConnect(RSQLite::SQLite(), dbname = "parcels.sqlite")
#create and connect to a sqlite database

dbWriteTable(con, "parcels", parcels, overwrite=TRUE)
#wrote the parcels dataframe into the database

dbWriteTable(con, "usecodes", UseCode, overwrite=TRUE)
#wrote the usecodes_raw dataframe into the database

dbWriteTable(con, "UseCode_Parcels", UseCodeParcels, overwrite=TRUE)
#wrote the combo dataframe into the database 

dbListTables(con)
#check the tables within the database.
```

## Analyze 

1. What are the 10 most-frequently-occuring land uses (in descending order)?

```{sql connection=con}
SELECT DISTINCT UseCode, 
CodeDesc AS "Land Use", 
count(*) AS "COUNT"
FROM UseCode_Parcels
GROUP BY  UseCode
ORDER BY "COUNT" DESC
LIMIT 10;
```
1. Single Family Residence: 73,915
2. Residential Rentals: 4,985
3. Sited inside M/H Park (Rented): 4,540
4. Vacant Land: 4,382
5. Condo: 3,791
6. Double Wide New Const. (after 7/1/80): 1,718
7. Apartments (5 or more units): 1,486
8. Rec. Area - 5-6 class, 1000-1300 sq.ft.: 1,423
9. Planned Unit Development: 1,124
10. Retail Stores, Single Story: 1,024

2. How many acres are in agricultural preserves? Assume agricultural preserves are indicated by non-NA values for AgPres.

```{sql connection=con}
SELECT ROUND(SUM(Acreage),2) 
AS "AG. Preserve Acreage"
FROM UseCode_Parcels
WHERE AgPres IS NOT NULL;
```
549563.4 acres on agricultural preserves

3. What is the mean net assessed value per acre of the entire county?
i.e., as if the entire county were a single parcel.

```{sql connection=con}
SELECT ROUND(SUM(Net_AV)/SUM(Acreage),3) AS "Mean Net Assessed Value/Acre"
FROM UseCode_Parcels
WHERE Acreage>0
AND Net_AV>0;
```
$60,833.24 mean net assessed value per acre for entire county.
#different value than assignment 1 taking out negative NET AV values this time

4. What is the total net assessed value of all non-taxable parcels?
Assume non-taxable parcels are indicated by non-NA values for NonTaxCode.

```{sql connection=con}
SELECT SUM(Net_AV) AS "Total net assessed value of all non-taxable parcels"
FROM UseCode_Parcels
WHERE NonTaxCode IS NOT NULL;
```
1,093,026,091 total net assessed value of all non-taxable parcels

5. What are the 10 largest property holders, by acreage? Use the complete mailing address for tax statements as a proxy for the parcel owner. Ignore parcels where either component of the address is NA.

```{sql connection=con}
SELECT M_Address1,M_Address2, SUM(Acreage) AS "Total Acreage"
FROM parcels
LEFT JOIN usecodes ON parcels.UseCode = usecodes.UseCode
WHERE M_Address1 IS NOT NULL
AND M_Address2 IS NOT NULL
GROUP BY M_Address1, M_Address2
ORDER BY "Total Acreage" DESC
LIMIT 10;
```
these slightly different than assignment 1

6. What are the 10 largest property holders, by net assessed value?

```{sql connection=con}
SELECT 
M_Address1 AS "Address 1",
M_Address2 AS "Address 2", 
SUM(Net_AV) AS "Total Value"
FROM parcels
LEFT JOIN usecodes ON parcels.UseCode = usecodes.UseCode
WHERE M_Address1 IS NOT NULL
AND M_Address2 IS NOT NULL
AND Net_AV > 0
GROUP BY M_Address1, M_Address2
ORDER BY "Total Value" DESC
LIMIT 10;
```
these slightly different than assignment 1